\bf On Misuses of the Kolmogorov--Smirnov Test for One-Sample Goodness-of-Fit

### 2024-06-13 manuscript/TAS-template.tex
put table label after caption

\documentclass[12pt, letterpaper]{article}
%\documentclass[12pt, letterpaper, titlepage]{article}

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = blue}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{setspace}

\usepackage[]{lineno}
\linenumbers*[1]
% %% patches to make lineno work better with amsmath
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
 \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
 \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
 \renewenvironment{#1}%
 {\linenomath\csname old#1\endcsname}%
 {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
 \patchAmsMathEnvironmentForLineno{#1}%
 \patchAmsMathEnvironmentForLineno{#1*}}%

\AtBeginDocument{%
 \patchBothAmsMathEnvironmentsForLineno{equation}%
 \patchBothAmsMathEnvironmentsForLineno{align}%
 \patchBothAmsMathEnvironmentsForLineno{flalign}%
 \patchBothAmsMathEnvironmentsForLineno{alignat}%
 \patchBothAmsMathEnvironmentsForLineno{gather}%
 \patchBothAmsMathEnvironmentsForLineno{multline}%
}

% control floats
\renewcommand\floatpagefraction{.9}
\renewcommand\topfraction{.9}
\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.1}
\setcounter{totalnumber}{50}
\setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}

\newcommand{\jy}[1]{\textcolor{blue}{JY: #1}}
\newcommand{\eds}[1]{\textcolor{red}{EDS: (#1)}}

% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{1}

%\title{On Misuses of the Kolmogorov--Smirnov Test for One-Sample Goodness-of-Fit}
%
%\author{Anthony Zeimbekakis\\
%%   \href{mailto:anthony.zeimbekakis@uconn.edu}
%% {\nolinkurl{anthony.zeimbekakis@uconn.edu}}\\
  %Elizabeth D.  Schifano\\
  %Jun Yan\\[1ex]
  %Department of Statistics, University of Connecticut\\
%}
%\date{}

\begin{document}
%\maketitle

\if0\blind
{
  \title{\bf On Misuses of the Kolmogorov--Smirnov Test for One-Sample 
	Goodness-of-Fit}
  \author{Anthony Zeimbekakis, %\\
%   \href{mailto:anthony.zeimbekakis@uconn.edu}
% {\nolinkurl{anthony.zeimbekakis@uconn.edu}}\\
  Elizabeth D. Schifano, %\\
  Jun Yan\\[1ex]
  Department of Statistics, University of Connecticut\\
}
\date{}
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf On Misuses of the Kolmogorov--Smirnov Test for One-Sample 
	Goodness-of-Fit}
\end{center}
  \bigskip
} \fi

\doublespace

\begin{abstract}
The Kolmogorov--Smirnov (KS) test is widely employed to assess the
goodness-of-fit of a hypothesized continuous distribution to a sample.
Despite its popularity, the test is frequently misused in the literature and
practice. While originally intended for independent, continuous data with
precisely specified hypothesized distributions, it is erroneously applied to
scenarios with dependent, discrete, or rounded data, with hypothesized
distributions requiring estimated parameters. For example, it has been
``discovered'' multiple times that the test is too conservative when the 
hypothesized distribution has parameters that need to be estimated.
We demonstrate misuses of the one-sample KS test in
three scenarios through simulation studies:
1) the hypothesized distribution has unspecified parameters;
2) the data are serially dependent; and
3) a combination of the first two scenarios.
For each scenario, we provide remedies for practical applications using
appropriate bootstrap approaches.
The whole demonstration can be used as hands-on education materials on both
goodness-of-fit tests and bootstrap.

\bigskip
\noindent{\sc Keywords}:
nonparametric bootstrap;
parametric bootstrap;
working dependence.
\end{abstract}

%\doublespace

\section{Introduction}
\label{sec:intro}

@@

The standard one-sample KS test applies to independent data with a continuous
hypothesized distribution that is completely specified. In practice, however, it
has often been applied without realizing that one or more of these assumptions
do not hold. For example, \citet{noether1963note} showed the conservativeness of
the KS test when applied to discontinuous distributions. The null distribution
of the KS statistic is no longer distribution-free and depends on the
hypothesized distribution~$F$. Computing the
exact and asymptotic distribution of $D_n$ is challenging. Fortunately, the null
distribution of the KS statistic for discontinuous distributions has been
efficiently addressed by \citet{dimitrova2020computing} with a
companion R package \textsf{KSgeneral}. Although a common misuse of the KS test, 
the issue with discontinuous data is not our focus.

@@

The standard KS test does not apply to stationary yet dependent data either. The
distribution of the KS statistic would have a higher variance for positively
dependent data than that derived when the data are independent because of a
smaller effective sample size. For example, for testing normality,
\citet{durilleul1992lack} demonstrate that a naive application of
the KS statistic is too liberal for medium-to-high positive serial dependence,
and that for negative dependence, the behavior is asymmetrical. For remedies,
\citet{weiss1978modification} provides a procedure that is applicable 
specifically for data
modeled by the second-order auto-regressive (AR) process where the AR parameters
are known. Serial dependence affects the validity of the two-sample KS test too.
In a recent paper on online diagnosis of performance variation in
high-performance computing systems, for example, the authors provided no details
about whether they accounted for serial dependence when applying the two-sample
KS test \citep{tuncer2019online}. After comparing various strategies for dealing
with serial dependence, \citet{lanzante2021testing} concludes that a test based
on Monte-Carlo simulations performed the best.
When additionally the hypothesized distribution contains unknown
parameters, the standard KS test becomes even further inapplicable.

The contribution of this paper is a demonstration of misuses of the one-sample
KS test in three scenarios and their remedies in practice. Contrary to the
assumptions within the statistics community that anyone familiar with the
relevant literature would not be ``misusing'' the KS test, there is still a
considerable journey ahead in educating students and practitioners from diverse
fields on the proper utilization of the KS test and similar tests. The scenarios
that we consider are where:
1) the hypothesized distribution has unspecified parameters;
2) the data are serially dependent; and
3) a combination of the first two scenarios.
In each scenario, the misuse is performed and the impacts are shown. Then, a
remedy is detailed and performed alongside the misuse to show its positive
effects. Specifically, unspecified parameters are handled by parametric
bootstrap; serial dependence is handled by introducing a working autoregressive
moving average (ARMA) model that preserves the serial dependence for a wide
range of dependence structures.
In order to set up the demonstrations, simulated data are used
throughout. The remedies are also performed on various families of
distributions. An R implementation of the proposed methods will be available
after the paper is published.

@@

The consequence of using the wrong null distribution for the KS statistic $D_n$
in~\eqref{eq:ks_fitted} can be illustrated through a simple simulation study. A
random sample $X_1, \ldots, X_n$ was generated from a normal
distribution with both mean and variance parameters set equal to 8 ($N(8,8)$), 
and sample size $n = 200$. The maximum likelihood estimates of the mean and 
variance were used as $\hat\theta_n$. The p-value of the test
statistic~$D_n$ in Equation~\eqref{eq:ks_fitted} was obtained by naively calling 
the \texttt{ks.test()} function in~R with the fitted normal distribution as the
hypothesized distribution. That is, the hypothesized distribution was
$N(\bar X, s^2)$ where $\bar X$ is the sample mean and $s^2$ is the sample
variance (with $n$ in the denominator). This experiment was repeated 1000 times
and the probability-probability plot (PP-plot) of the 1000 p-values are 
displayed in the top left plot of Figure~\ref{fig:pp_f}. If the test were
valid, the p-values would be uniformly distributed in the unit
interval and the points in the PP-plot would fall along the diagonal line. 
Clearly, the distribution of the 1000 p-values is very different from what
one would expect to see from 1000 draws from the standard uniform
distribution.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=.5\textwidth]{pp_f.pdf}
  \caption{PP-plots for the p-values obtained from KS tests with unspecified
    parameters with sample size $n = 200$ based on 1000 replicates. The data
    generating models were $N(8, 8)$ and $\Gamma(8, 1)$.
  }
  \label{fig:pp_f}
\end{figure}

We similarly considered random samples $X_1, \ldots, X_n$ 
generated from a gamma distribution with shape and scale parameters 8 and 1, 
respectively ($\Gamma(8,1)$), and sample size $n=200$. These parameter values 
were selected so that the mean and variance of the normal and gamma distributions 
in the simulations are the same. Likewise, the p-value of the test 
statistic~$D_n$ in~\eqref{eq:ks_fitted} was obtained by naively calling the
\texttt{ks.test()} function in~R with the fitted gamma distribution as the
hypothesized distribution using the shape and scale maximum likelihood 
estimates. The PP-plot of the p-values from 1000 replications of this 
experiment are displayed in the bottom left plot of Figure~\ref{fig:pp_f}. The
p-values are again non-uniformly distributed due to the use of the 
estimated parameters when specifying the null distribution.

@@

Figure~\ref{fig:pp_ssf} shows the PP-plots resulting p-values of the
procedure performed on data generated from same simulation settings as in
Section~\ref{sec:dependence}. For comparison, we also displayed the PP-plot of
a naive bootstrap procedure that adjusts for the unspecified parameters as in
Section~\ref{sec:fitted} but not for the serial dependence. 
The super naive approach that does not adjust for fitted parameters nor serial
dependence would have double sin, so the results are not reported. 
The p-values from
naively fitting parameters but not adjusting for dependence clearly deviates
from the standard uniform distribution except in the MA(1) case, where the
autocorrelation is nonzero with a short memory of only lag~1.
Similar to Section~\ref{sec:dependence}, the working normal copula approach
gives p-values with the desired standard uniform distribution regardless of the
true dependence.

@@

\begin{table}[tbp]
  \label{tab:power}
  \caption{Power of rejecting the null hypothesis of normal and gamma
    distribution at significance level 0.05 when the true distribution was
    $\Gamma(8, 1)$ and $N(8, 8)$, respectively. }
\centering
\begin{tabular}{ll cccc}
  \toprule
  &  & \multicolumn{2}{c}{$H_0:$ normal} &
  \multicolumn{2}{c}{$H_0$: gamma}\\
  \cmidrule(lr){3-4}\cmidrule(lr){5-6}
  Dependence & Method & $n = 100$ & $n = 200$ & $n = 100$ & $n = 200$\\
  \midrule
ARMA & naive & 0.404 & 0.656 & 0.515 & 0.779 \\ 
 & copula & 0.330 & 0.556 & 0.442 & 0.722 \\ 
AR & naive & 0.411 & 0.655 & 0.455 & 0.705 \\ 
& copula & 0.261 & 0.448 & 0.313 & 0.544 \\ 
MA & naive & 0.374 & 0.728 & 0.514 & 0.817 \\ 
& copula & 0.343 & 0.693 & 0.487 & 0.797 \\ 
   \bottomrule
\end{tabular}
\end{table}

\section{Classroom Implementation}

% Since this is a Teacher's Corner submission, I was hoping to read about how
% these ideas have been presented in specific courses and also how students
% reacted to this. What background did the students have? What level of detail was
% given to them? That is, it is one thing to say "use the bootstrap" and another
% thing to present to students, e.g., a Markdown file with all necessary code
% included. Was that done? Were students invited to conduct their own simulations?
% Etc.

\if0\blind
{
% Graduate level
The misuses of one-sample KS test can be easily incorporated into graduate level
classroom teaching for courses that cover parametric bootstrap. For example, at
the University of Connecticut (UConn), JY has always included a demonstration of
the misuse under unspecified parameters in STAT 5361, Statistical Computing. The
catalog description of this course includes: ``Use of computing for statistical
problems; obtaining features of distributions, fitting models and implementing
inference. Basic numerical methods, nonlinear statistical methods, numerical
integration, modern simulation methods.'' Bootstrap is covered under ``modern
simulation methods.''  The course is open to graduate students in Statistics,
others with permission. Detailed R code and graphical illustrations available in
the open-access course notes elucidate the distinctions between employing null
hypotheses with and without defined parameters (accessible at:
\url{https://jun-yan.github.io/stat-5361/boot.html#parametric-bootstrap}).
Students were able to follow the demonstration closely, gaining a clear
understanding of the implications of using estimated parameters within test
statistics, and learning about the efficacy of the parametric bootstrap method
as a corrective measure.

Supplementary materials accompanying this paper present R code that simplifies
the illustration of the three common misuses of the one-sample KS test in
classrooms, whether during lectures or as part of homework assignments.
Addressing the misuses in contexts of serial dependency necessitates familiarity
with ARMA models in time series analysis---a prerequisite achievable by most
statistics graduate students. While incorporating copulas to model dependency
structures extends beyond conventional statistics curricula, it presents an
opportunity for advanced exploration in courses focused on statistical computing
or multivariate analysis.

% Undergraduate level (could mention AZ's undergraduate thesis)
At the undergraduate level, so far only the misuses of one-sample KS test have 
been presented and demonstrated in a classroom setting. At UConn, ES 
incorporated this material in STAT 3215, Applied Linear Regression in Data 
Science, as the one-sample KS test is used in residual diagnostics for testing 
whether the externally studentized residuals can be viewed as realizations of a 
standard normal distribution for large samples. %The catalog description of this 
%course states the following topics: ``Applied multiple 
%linear regression analysis in data science, with an emphasis on modern 
%statistical regression methods: simple linear regression and correlation 
%analysis, multiple linear regression, analysis of variance, goodness of fit, 
%comparing regression models through partial and sequential F tests, dummy 
%variables, regression assumptions and diagnostics, model selection and penalized 
%regression, prediction and model validation, principles of design of experiments, 
%one-way and two-way analysis of variance.'' 
Students in this applied linear regression course are required
to have taken either two semesters of introductory statistics or a 
Calculus~2-based statistical methods course as a prerequisite. In the KS 
assignment, students were introduced to a more detailed description of the KS 
test, and were provided R code and output for a simulation study that showed  
KS test performance (in terms of the distribution of $p$-values) under a 
scenario where all KS test assumptions were met (independent, continuous data 
with precisely specified hypothesized distributions). When the scenario was 
modified to use estimated parameters instead, the students were asked modify the 
previous simulation code and show the resulting distribution of $p$-values. The 
assignment is included in the Supplemental Materials. Students were able to
properly simulate and correctly produce results indicating the conservative 
nature of the test under these conditions.

While students in this course are exposed to the bootstrap in the 
context of linear regression models (specifically, case resampling and residual 
bootstrap), they are not in general familiar with serially dependency. Thus, the 
remedies and more complicated serially dependent settings were not presented in 
the undergraduate applied linear regression course. Serial dependence is covered 
in other undergraduate courses, and in fact, the contents of this manuscript 
were the basis of AZ's undergraduate honors thesis. Consequently, all 
statistical concepts and methods employed within this paper are certainly within 
reach of many UConn undergraduate Statistics and Statistical Data Science 
majors. 
} \fi

\if1\blind
{
% Graduate level
The misuses of one-sample KS test can be easily incorporated into graduate level
classroom teaching for courses that cover parametric bootstrap. For example, at
the University of XX, YY has always included a demonstration of the misuse under 
unspecified parameters in STAT $\#\#\#\#$, XXXX. The
catalog description of this course includes: ``Use of computing for statistical
problems; obtaining features of distributions, fitting models and implementing
inference. Basic numerical methods, nonlinear statistical methods, numerical
integration, modern simulation methods.'' Bootstrap is covered under ``modern
simulation methods.''  The course is open to graduate students in Statistics,
others with permission. Detailed R code and graphical illustrations available in
the open-access course notes elucidate the distinctions between employing null
hypotheses with and without defined parameters (accessible at: 
\textit{author YY's github website; material printed to pdf for blinded review}).
%\url{https://jun-yan.github.io/stat-5361/boot.html#parametric-bootstrap}).
Students were able to follow the demonstration closely, gaining a clear
understanding of the implications of using estimated parameters within test
statistics, and learning about the efficacy of the parametric bootstrap method
as a corrective measure.

Supplementary materials accompanying this paper present R code that simplifies
the illustration of the three common misuses of the one-sample KS test in
classrooms, whether during lectures or as part of homework assignments.
Addressing the misuses in contexts of serial dependency necessitates familiarity
with ARMA models in time series analysis---a prerequisite achievable by most
statistics graduate students. While incorporating copulas to model dependency
structures extends beyond conventional statistics curricula, it presents an
opportunity for advanced exploration in courses focused on statistical computing
or multivariate analysis.

% Undergraduate level (could mention AZ's undergraduate thesis)
At the undergraduate level, so far only the misuses of one-sample KS test have 
been presented and demonstrated in a classroom setting. At XX, ZZ 
incorporated this material in STAT $\#\#\#\#$, XXXX, as the one-sample KS test 
is used in residual diagnostics for testing 
whether the externally studentized residuals can be viewed as realizations of a 
standard normal distribution for large samples. %The catalog description of this 
%course states the following topics: ``Applied multiple 
%linear regression analysis in data science, with an emphasis on modern 
%statistical regression methods: simple linear regression and correlation 
%analysis, multiple linear regression, analysis of variance, goodness of fit, 
%comparing regression models through partial and sequential F tests, dummy 
%variables, regression assumptions and diagnostics, model selection and penalized 
%regression, prediction and model validation, principles of design of experiments, 
%one-way and two-way analysis of variance.'' 
Students in this applied linear regression course are required
to have taken either two semesters of introductory statistics or a 
Calculus~2-based statistical methods course as a prerequisite. In the KS 
assignment, students were introduced to a more detailed description of the KS 
test, and were provided R code and output for a simulation study that showed  
KS test performance (in terms of the distribution of $p$-values) under a 
scenario where all KS test assumptions were met (independent, continuous data 
with precisely specified hypothesized distributions). When the scenario was 
modified to use estimated parameters instead, the students were asked modify the 
previous simulation code and show the resulting distribution of $p$-values. The 
assignment is included in the Supplemental Materials. Students were able to
properly simulate and correctly produce results indicating the conservative 
nature of the test under these conditions.

While students in this course are exposed to the bootstrap in the 
context of linear regression models (specifically, case resampling and residual 
bootstrap), they are not in general familiar with serially dependency. Thus, the 
remedies and more complicated serially dependent settings were not presented in 
the undergraduate applied linear regression course. Serial dependence is covered 
in other undergraduate courses, and in fact, the contents of this manuscript 
were the basis of WW's undergraduate honors thesis. Consequently, all 
statistical concepts and methods employed within this paper are certainly within 
reach of many XX undergraduate Statistics and Data Science majors. 
} \fi

\section{Concluding Remarks}
\label{sec:conclusion}

The three kinds of misuses of the KS test for one-sample goodness-of-fit that we
demonstrated here are commonly seen in practice. With the aid of computing tools
such as R, they can be easily demonstrated in classroom teaching. Practitioners
need to be aware of the assumptions of the standard one-sample KS test that the
hypothesized distribution is completely specified and that the data are
independent. When these assumptions are not met, naive applications of the test
are no longer accurate and remedies must be performed. In the
case of fitted parameters, parametric and nonparametric bootstrap can restore
the size of the test. The nonparametric version has long been developed too but
is lesser known \citep{babu2004goodness}. In the case of serially dependent data,
our semiparametric bootstrap procedure with a working dependence structure
controlled by an automatically selected ARMA model restores the size of the test
for a wide range of dependence structures.
When both assumptions are violated, i.e., where the data has serial 
dependence and the parameters must be fitted, a combination of the two
individual remedies shows good results. The misuses of the KS test also apply
to other goodness-of-fit tests based on the empirical distribution
function, such as the Anderson--Darling test or the Cramér-—von Mise test, which
are known to be superior to KS in certain respects \citep{stephens2017tests}.
Similar bootstrap remedies can be developed for these tests.

The semiparametric bootstrap procedure with a working ARMA dependence structure
is not a complete solution and has limitations. When the true serial dependence
can be well approximated by an ARMA model, the model selection process could be
replaced with
autoregressive sieve, i.e., a sequence of autoregressive models the order of 
which increases
simultaneously with the sample size although at a suitably slower rate 
\citep[e.g.,][]{psaradakis2017distance, psaradakis2020normality}.
When the true serial dependence
structure cannot be well approximated by an ARMA model with normally distributed
error terms, the remedy may not restore the size of the test. The working
dependence structure is essentially the normal copula uniquely determined by the
ARMA model with normal errors. Some copulas are quite different from the normal
copula. For example, extreme value copulas have tail dependence, which is not
provided by normal copulas.
Using a normal copula to approximate an extreme value copula may not capture the
true serial dependence structure that is needed to restore the size of the
test. A block version of the nonparametric bootstrap \citep{babu2004goodness},
which avoids specification of a working dependence structure, would be more
desirable.

Statisticians living in
an ivory tower tend to doubt that anyone would be misusing the KS test as we
illustrated given that such situations are very well-known in the statistics
literature. It is our experience, however, that much effort is needed in
propagating many knowns in the statistics community to the real wild world
through innovative and impactful education. Such effort may be more valuable
than developing more advanced methods that few people would use in practice.

\section*{Data Availability Statement}
An R implementation of the proposed methods as well as teaching materials are  
included in the Supplementary Materials.

@@

\section{Appendix: Nonparametric Bootstrap for KS Test with Fitted Parameters}

@@

\bibliographystyle{asa}
\bibliography{citations}

\end{document}
%%% LocalWords: nonparametric semiparametric autocorrelation ARMA
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% ispell-personal-dictionary: ".aspell.en.pws"
%%% fill-column: 80
%%% eval: (auto-fill-mode 1)
%%% End:

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{natbib}
%\usepackage{url} % not crucial - just used below for the URL 
\usepackage{booktabs}
\usepackage{hyperref}
\hypersetup{colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = blue}
\usepackage{enumitem}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\begin{document}

%\bibliographystyle{natbib}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf On Misuses of the Kolmogorov--Smirnov Test for One-Sample 
	Goodness-of-Fit}
  \author{Anthony Zeimbekakis\\
%   \href{mailto:anthony.zeimbekakis@uconn.edu}
% {\nolinkurl{anthony.zeimbekakis@uconn.edu}}\\
  Elizabeth D. Schifano\\
  Jun Yan\\[1ex]
  Department of Statistics, University of Connecticut\\
}
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf On Misuses of the Kolmogorov--Smirnov Test for One-Sample 
	Goodness-of-Fit}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
The Kolmogorov--Smirnov (KS) test is one of the most popular goodness-of-fit
tests for comparing a sample with a hypothesized continuous distribution.
Nevertheless, it has often been misused. The standard one-sample KS test applies
to independent, continuous data with a hypothesized distribution that is
completely specified. It is not uncommon, however, to see in the literature that
it was applied to dependent, discrete, or rounded data, with hypothesized
distributions containing estimated parameters. For example, it has been
``discovered'' multiple times that the test is too conservative when the 
hypothesized distribution has parameters that need to be estimated.
We demonstrate misuses of the one-sample KS test in
three scenarios through simulation studies:
1) the hypothesized distribution has unspecified parameters;
2) the data are serially dependent; and
3) a combination of the first two scenarios.
For each scenario, we provide remedies for practical applications.
\end{abstract}

\noindent%
{\it Keywords:}  nonparametric bootstrap; parametric bootstrap;
working dependence.
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!
\section{Introduction}
\label{sec:intro}

@@

The standard one-sample KS test applies to independent data with a continuous
hypothesized distribution that is completely specified. In practice, however, it
has often been applied without realizing that one or more of these assumptions
do not hold. For example, \citet{noether1963note} showed the conservativeness of
the KS test when applied to discontinuous distributions. The null distribution
of the KS statistic is no longer distribution free. Computing the
exact and asymptotic distribution of $D_n$ is challenging. Fortunately, the null
distribution of the KS statistic for discontinuous distributions has been
efficiently addressed by \citet{dimitrova2020computing} with a
companion R package \textsf{KSgeneral}. Although a common misuse of the KS test, 
the issue with discontinuous data is not our focus.

@@

The standard KS test does not apply to stationary yet dependent data either. The
distribution of the KS statistic would have a higher variance for positively
dependent data than that derived when the data are independent because of a
smaller effective sample size. For example, for testing normality,
\citet{durilleul1992lack} demonstrates that a naive application of
the KS statistic is too liberal for medium-to-high positive serial dependence,
and that for negative dependence, the behavior is asymmetrical. For remedies,
\citet{weiss1978modification} provides a procedure that is applicable 
specifically for data
modeled by the second-order auto-regressive (AR) process where the AR parameters
are known. \citet{lanzante2021testing} tests various strategies for dealing with
temporal
dependence and concludes that a test based on Monte-Carlo simulations performed
the best. When additionally the hypothesized distribution contains unknown
parameters, the standard KS test becomes even further inapplicable.

The contribution of this paper is a demonstration of misuses of the one-sample
KS test in three scenarios and their remedies in practice. The scenarios are
where:
1) the hypothesized distribution has unspecified parameters;
2) the data are serially dependent; and
3) a combination of the first two scenarios.
In each scenario, the misuse is performed and the impacts are shown. Then, a
remedy is detailed and performed alongside the misuse to show its positive
effects. Specifically, unspecified parameters are handled by parametric
bootstrap; serial dependence is handled by introducing a working autoregressive
moving average (ARMA) model that preserves the serial dependence for a wide
range of dependence structures.
In order to set up the demonstrations, simulated data are used
throughout. The remedies are also performed on various families of
distributions. An R implementation of the proposed methods will be available
after the paper is published.

@@

The consequence of using the wrong null distribution for the KS statistic $D_n$
in~\eqref{eq:ks_fitted} can be illustrated through a simple simulation study. A
random sample $X_1, \ldots, X_n$ was generated from a normal
distribution with both mean and variance parameters set equal to 8 ($N(8,8)$), 
and sample size $n = 200$. The maximum likelihood estimates of the mean and 
variance were used as $\hat\theta_n$. The p-value of the test
statistic~$D_n$ in Equation~\eqref{eq:ks_fitted} was obtained by naively calling 
the \texttt{ks.test()} function in~R with the fitted normal distribution as the
hypothesized distribution. That is, the hypothesized distribution was
$N(\bar X, s^2)$ where $\bar X$ is the sample mean and $s^2$ is the sample
variance (with $n$ in the denominator). This experiment was repeated 1000 times
and the probability-probability plot (PP-plot) of the 1000 p-values is 
displayed in the top left plot of Figure~\ref{fig:pp_f}. If the test were
valid, the p-values would be uniformly distributed in the unit
interval and the points in the PP-plot would fall along the diagonal line. 
Clearly, the distribution of the 1000 p-values is very different from what
one would expect to see from 1000 draws from the standard uniform
distribution.

\begin{figure}
  \centering
  \includegraphics[width=.5\textwidth]{pp_f.pdf}
  \caption{PP-plots for the p-values obtained from KS tests with unspecified
    parameters with sample size $n = 200$ based on 1000 replicates. The data
    generating models were $N(8, 8)$ and $\Gamma(8, 1)$.
 }
 \label{fig:pp_f}
\end{figure}

We similarly considered random samples $X_1, \ldots, X_n$ 
generated from a gamma distribution with shape and scale parameters 8 and 1, 
respectively ($\Gamma(8,1)$), and sample size $n=200$. These parameter values 
were selected so that the mean and variance of the normal and gamma distributions 
were the same in the simulations. Likewise, the p-value of the test 
statistic~$D_n$ in~\eqref{eq:ks_fitted} was obtained by naively calling the
\texttt{ks.test()} function in~R with the fitted gamma distribution as the
hypothesized distribution using the shape and scale maximum likelihood 
estimates. The PP-plot of the p-values from 1000 replications of this 
experiment is displayed in the bottom left plot of Figure~\ref{fig:pp_f}. The
p-values are again non-uniformly distributed due to the use of the 
estimated parameters when specifying the null distribution.

@@

Figure~\ref{fig:pp_ssf} shows the resulting PP-plots of p-values from the
procedure performed on data generated from the same simulation settings as in
Section~\ref{sec:dependence}. For comparison, we also display the PP-plots from
a naive bootstrap procedure that adjusts for the unspecified parameters as in
Section~\ref{sec:fitted} but not for the serial dependence. 
The super naive approach that does not adjust for fitted parameters nor serial
dependence would have double sin, so the results are not reported. 
The p-values from
naively fitting parameters but not adjusting for dependence clearly deviate
from the standard uniform distribution except in the MA(1) case, where the
autocorrelation is nonzero with a short memory of only lag~1.
Similar to Section~\ref{sec:dependence}, the working normal copula approach
gives p-values with the desired standard uniform distribution regardless of the
true dependence.

\begin{table}[tbp]
  \caption{Power of rejecting the null hypothesis of normal and gamma
    distribution at significance level 0.05 when the true distribution was
    $\Gamma(8, 1)$ and $N(8, 8)$, respectively. }
  \label{tab:power}
\centering
\begin{tabular}{ll cccc}
  \toprule
  &  & \multicolumn{2}{c}{$H_0:$ normal} &
  \multicolumn{2}{c}{$H_0$: gamma}\\
  \cmidrule(lr){3-4}\cmidrule(lr){5-6}
  Dependence & Method & $n = 100$ & $n = 200$ & $n = 100$ & $n = 200$\\
  \midrule
ARMA & naive & 0.404 & 0.656 & 0.515 & 0.779 \\ 
 & copula & 0.330 & 0.556 & 0.442 & 0.722 \\ 
AR & naive & 0.411 & 0.655 & 0.455 & 0.705 \\ 
& copula & 0.261 & 0.448 & 0.313 & 0.544 \\ 
MA & naive & 0.374 & 0.728 & 0.514 & 0.817 \\ 
& copula & 0.343 & 0.693 & 0.487 & 0.797 \\ 
   \bottomrule
\end{tabular}
\end{table}

@@

\section{Conclusion}
\label{sec:conclusion}

The three kinds of misuses of the KS test for one-sample goodness-of-fit that we
demonstrated here are commonly seen in practice. With the aid of computing tools
such as R, they can be easily demonstrated in classroom teaching. Practitioners
need to be aware of the assumptions of the standard one-sample KS test that the
hypothesized distribution is completely specified and that the data are
independent. When these assumptions are not met, naive applications of the test
are no longer accurate and remedies must be performed. In the
case of fitted parameters, parametric and nonparametric bootstrap can restore
the size of the test. The nonparametric version has long been developed too but
is lesser known \citep{babu2004goodness}. In the case of serially dependent data,
our semiparametric bootstrap procedure with a working dependence structure
controlled by an automatically selected ARMA model restores the size of the test
for a wide range of dependence structures.
When both assumptions are violated, i.e., where the data has serial 
dependence and the parameters must be fitted, a combination of the two
individual remedies shows good results.

The semiparametric bootstrap procedure with a working ARMA dependence structure
is not a complete solution and has limitations. When the true serial dependence
structure cannot be well approximated by an ARMA model with normally distributed
error terms, the remedy may not restore the size of the test. The working
dependence structure is essentially the normal copula uniquely determined by the
ARMA model with normal errors. Some copulas are quite different from the normal
copula. For example, extreme value copulas have tail dependence, which is not
provided by normal copulas.
Using a normal copula to approximate an extreme value copula may not capture the
true serial dependence structure that is needed to restore the size of the
test. A block version of the nonparametric bootstrap \citep{babu2004goodness},
which avoids specification of a working dependence structure, would be more
desirable.

@@

\section{Nonparametric Bootstrap for KS Test with Fitted Parameters}

@@

\bibliographystyle{asa}
\bibliography{citations}
\end{document}
